---
title: "Customer Analytics Project"
output: 
  flexdashboard::flex_dashboard:
    theme: journal
    
    orientation: columns
    vertical_layout: fill
    
---
<style>                     
.navbar {
  background-color:green;
  border-color:blue;
}
.navbar-brand {
color:black!important;
}
</style>           
---

```{r setup, include=FALSE}
library(flexdashboard)
library(data.table)
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
library(DT)
library(ggthemes)
library(plyr)
library(readr)
library(arules)
library(arulesViz)
library(methods)
library(treemap)
library(visNetwork)
library(igraph)

orders <- fread('data/orders.csv')
products <- fread('data/products.csv')
order_products <- fread('data/order_products__train.csv')
order_products_prior <- fread('data/order_products__prior.csv')
aisles <- fread('data/aisles.csv')
departments <- fread('data/departments.csv')


```

Exploratory Data Analysis
=====================================  
    
Column {.tabset}
-------------------------------------
   
### Dataset OverView

Lets first have a look at provided datasets:

#### orders

This file gives a list of all orders we have in the dataset. 1 row per order. 

```{r, result='asis'}
kable(head(orders,5))
glimpse(orders)
```


#### order_products_train

This file gives us information about which products (product_id) were ordered. It also contains information of the order (add_to_cart_order) in which the products were put into the cart and information of whether this product is a re-order(1) or not(0).


```{r}
kable(head(order_products,5))
glimpse(order_products)
```

#### products

This file contains the names of the products with their corresponding product_id. 

```{r}
kable(head(products,5))
glimpse(products)
```

#### order_products_prior

This file is structurally the same as the other_products_train.csv. 

```{r, result='asis'}
kable(head(order_products_prior,5))
glimpse(order_products_prior)
```


#### aisles

This file contains the different aisles.

```{r, result='asis'}
kable(head(aisles,5))
glimpse(aisles)
```

#### departments

```{r, result='asis'}
kable(head(departments,5))
glimpse(departments)
```

```{r message=FALSE, warning=FALSE}
orders <- orders %>% mutate(order_hour_of_day = as.numeric(order_hour_of_day), eval_set = as.factor(eval_set))
products <- products %>% mutate(product_name = as.factor(product_name))
aisles <- aisles %>% mutate(aisle = as.factor(aisle))
departments <- departments %>% mutate(department = as.factor(department))
```


### Order Frequency

When people buy groceries online. 


#### Hour of Day
Most people order between 9:00 AM to 6:00 PM in the evening.
```{r, fig.width=4.5,fig.height=3.5}
orders %>% 
  ggplot(aes(x=order_hour_of_day,color="coralred",fill="blue")) + 
  geom_histogram(stat="count")+
  theme_hc()+ scale_colour_hc()

```

#### Day of Week
Sunday and Monday are the days when people order most on Instacart

```{r, fig.width=4.5,fig.height=3.5}

orders %>% 
  ggplot(aes(x=order_dow,color="coralred",fill="coralred")) + 
  geom_histogram(stat="count")+
  theme_hc()+ scale_colour_hc()
```

#### Re-order analysis
find two categories of people! One that reorders monthly other who does weekly. This is based on the peaks formed at 30th day and 7th day. 
```{r, fig.width=7.5,fig.height=4.5}
library(plyr)
library(dplyr)
Reordering_Gap<-count(orders,'days_since_prior_order')%>%arrange(desc(freq))%>%mutate(Percent_orders=round(freq*100/nrow(orders)),2)

#Inference: 11 % of the time people reorder monthly(after 30 days), and 9 % of the time weekly. This shows there is a section of people who refill their groceries every month and other who refills every week. Frequency of NA represents total number of unique users and its their first order.

#Visualizing reordering Gap

Reordering_Gap_plot<-ggplot(orders,aes(x=days_since_prior_order))+
  geom_histogram(aes(fill=..count..),binwidth=1)+
  scale_x_continuous(name = "Days Since Prior Order",breaks = seq(0, 30, 1))+
  scale_y_continuous(name = "Frequency of Orders",breaks=seq(0,1000000,100000))+
  ggtitle("When customers order")+
  labs(x="Days Since Prior Order")+
  theme_update()
Reordering_Gap_plot
```


### Department

#### Which department is most commonly purchased by Day of Week (Top 5 Departments)
```{r, fig.width=6,fig.height=5}

products_departments <-
  merge(products, departments, by = "department_id")

## Creating initial data frame - nothing aggregated
sku_level_orders <- left_join(
  right_join(orders, order_products[, c("order_id", "product_id")],
             by = "order_id"),
  products_departments,
  by = "product_id"
)

## Most popular departments by day of week (all departments ranked)
department_purchases_dow <-
  aggregate(product_id ~ department + order_dow, sku_level_orders, FUN = length)
names(department_purchases_dow) <-
  c("department", "order_dow", "num_of_orders")

## Creating vector corresponding to day of week name order was placed 
department_purchases_dow$day_week_name <-
  ifelse(
    department_purchases_dow$order_dow == 0,
    "Sunday",
    ifelse(
      department_purchases_dow$order_dow == 1,
      "Monday",
      ifelse(
        department_purchases_dow$order_dow == 2,
        "Tuesday",
        ifelse(
          department_purchases_dow$order_dow == 3,
          "Wednesday",
          ifelse(
            department_purchases_dow$order_dow == 4,
            "Thursday",
            ifelse(
              department_purchases_dow$order_dow == 5,
              "Friday",
              ifelse(department_purchases_dow$order_dow == 6, "Saturday", "")
            )
          )
        )
      )
    )
  )


## Creating list of week days to run through for loop
day_of_week <- unique(department_purchases_dow$day_week_name)


## Temporary dataframes used in for loop
temp_df <- ""
temp_df <- as.data.frame(temp_df)
x_temp <- ""
x_temp <- as.data.frame(x_temp)
department_dow_top5 <- ""
department_dow_top5 <- as.data.frame(department_dow_top5)

# Top N Number of departments to graph
n = 5
for (each_day in day_of_week) {
  temp_df <-
    department_purchases_dow[department_purchases_dow$day_week_name %in% each_day, ]
  x_temp <- temp_df[order(-temp_df$num_of_orders),][1:n, ]
  department_dow_top5 <- rbind.fill(department_dow_top5, x_temp)
  department_dow_top5 <-
    department_dow_top5[!is.na(department_dow_top5$department), c("department", "order_dow", "num_of_orders", "day_week_name")]
  
}


## Reording day names for ggplot
department_dow_top5$day_ordered = factor(
  department_dow_top5$day_week_name,
  levels = c(
    "Sunday",
    "Monday",
    "Tuesday",
    "Wednesday",
    "Thursday",
    "Friday",
    "Saturday"
  )
)


### Creating and storing ggplot 
top5_department_by_dow <-
  ggplot(department_dow_top5,
         aes(reorder(department,-num_of_orders), num_of_orders, fill = department)) +
  geom_bar(
    stat = "identity",
    position = position_dodge(0.5),
    colour = "black",
    width = .8
  ) +
  facet_grid(. ~ day_ordered) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),plot.title = element_text(hjust = 0.5)
        ) +
  guides(fill = guide_legend(title ="Department Color")) +
  ggtitle("Top 5 Department in # of Orders by DOW") +
  labs(x = "Deparment",  y = "Number of Orders")

## Printing ggplot
top5_department_by_dow

```

#### What are the Top 10 Departments being purchased during Customer's first order?

```{r, fig.width=6,fig.height=4.5}
first_order_ID <- orders[orders$order_number %in% "1", c("order_id")] 

first_order_products <- left_join(
  left_join(order_products, first_order_ID, by = c("order_id")), products_departments, by = "product_id"
)

first_order_products$order_id <- as.character(first_order_products$order_id)
first_order_products$product_count <- 1
top_departments_firstOrder <-
  aggregate(order_id ~ department, first_order_products, length)

names(top_departments_firstOrder) <-
  c("department", "First_Order_Frequency")

## Sorting by department frequency and keeping only top 10 departments purchased during first order
top_departments_firstOrder_top10 <-
  top_departments_firstOrder[order(-top_departments_firstOrder$First_Order_Frequency), ][1:10, ]

## Plotting top 10 departments purchased on a customer's first order
ggplot(top_departments_firstOrder_top10,
       aes(
         reorder(department,-First_Order_Frequency),
         First_Order_Frequency
       )) +
  geom_bar(stat = "identity",
           fill = "chartreuse2",
           colour = "black") +
  stat_summary(
    aes(label = ..y..),
    geom = "text",
    vjust = +1.2,
    size = 4
  ) +
  labs(title = "Top 10 Departments Purchased on Customer's First Order",
       x = "Department", y = "# of Purchases during First Order") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Items
    
#### How many items does a customer order have ?
```{r, fig.width=6,fig.height=4.5}
total_order_size <- order_products_prior %>% 
    select(add_to_cart_order) %>%
        group_by(add_to_cart_order) %>%
            dplyr::summarise(count=n())
#head(total_order_size)
ggplot(total_order_size) + aes(add_to_cart_order,count,color="coralred",fill="darkgreen") + geom_line() + 
    ggtitle('Number of Items in Prior set Orders') + labs(y= "Number of Orders", x = "Items in Order")
```

#### Items sold most often (top10) 

```{r fig.height=5.5}
tmp <- order_products %>% 
  group_by(product_id) %>% 
  dplyr::summarize(count = n()) %>% 
  top_n(10, wt = count) %>%
  left_join(select(products,product_id,product_name),by="product_id") %>%
  arrange(desc(count)) 
#kable(tmp)

tmp %>% 
  ggplot(aes(x=reorder(product_name,-count), y=count,color="coralred",fill="darkgreen"))+
  geom_bar(stat="identity")+
  theme(axis.text.x=element_text(angle=90, hjust=1),axis.title.x = element_blank())

```
<br/>
Banana is most purchase item followed Organic Strawberries and Baby Spinach.



### Product Portfolio

#### Visualizing the structure of instacarts product portfolio

```{r, fig.width=10,fig.height=7}

tmp <- products %>% group_by(department_id, aisle_id) %>% dplyr::summarize(n=n())
tmp <- tmp %>% left_join(departments,by="department_id")
tmp <- tmp %>% left_join(aisles,by="aisle_id")

tmp2<-order_products %>% 
  group_by(product_id) %>% 
  dplyr::summarize(count=n()) %>% 
  left_join(products,by="product_id") %>% 
  ungroup() %>% 
  group_by(department_id,aisle_id) %>% 
  dplyr::summarize(sumcount = sum(count)) %>% 
  left_join(tmp, by = c("department_id", "aisle_id")) %>% 
  mutate(onesize = 1)

treemap(tmp2,index=c("department","aisle"),vSize="onesize",vColor="department",palette="Set3",title="",sortID="-sumcount", border.col="#FFFFFF",type="categorical", fontsize.legend = 0,bg.labels = "#FFFFFF")

```

Market Basket Analysis
=====================================  
    
Column {.tabset}
-------------------------------------

### Rules->For Products

#### Rules generated with Support=0.01% and Confidence=50%

```{r message=FALSE, warning=FALSE,include=FALSE}

library(XML)
library(pmml)
rules1 = read.PMML("data/rules1.xml")

```

```{r,fig.width=15,fig.height=5}
rules_dt1 <- data.table( lhs = labels( lhs(rules1) ), 
                        rhs = labels( rhs(rules1) ), 
                        quality(rules1) )[ order(-lift), ]
kable(head(rules_dt1,10))
glimpse(rules_dt1)


```

### Rules- Scatter Plot (For products)

```{r,fig.width=8,fig.height=8}
library(data.table)
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
library(DT)
library(arules)
library(arulesViz)
library(plotly)
library(visNetwork)
library(igraph)

plotly_arules(rules1)

```


```{r message=FALSE, warning=FALSE,include=FALSE}
    subrules2 <- head(sort(rules1, by="confidence"),10)
    ig <- plot( subrules2, method="graph", control=list(type="items") )
    ig_df <- get.data.frame( ig, what = "both" )
    nodesv <- data.frame(
      id = ig_df$vertices$name
      ,value = ig_df$vertices$support # could change to lift or confidence
      ,title = ifelse(ig_df$vertices$label == "",ig_df$vertices$name, ig_df$vertices$label)
      ,ig_df$vertices
    ) 
    
   
    ```
      
      
### Rules- Network Graph Visual ( For products)
    ```{r,fig.width=10,fig.height=10}
      ### Network graph visualization Part 2
    
    
    visNetwork(nodes = nodesv, edges =ig_df$edges, width = 900, height = 700) %>%
      visNodes(size = 10) %>%
      visLegend() %>%
      visEdges(smooth = FALSE) %>%
      visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) %>%
      visInteraction(navigationButtons = TRUE) %>%
      visEdges(arrows = 'from') %>%
      visPhysics(
        solver = "barnesHut",
        maxVelocity = 35,
        forceAtlas2Based = list(gravitationalConstant = -6000)
      )
    

```


### Rules->For Aisles

#### Rules generated with Support=0.7% and Confidence=40%

```{r message=FALSE, warning=FALSE,include=FALSE}

library(XML)
library(pmml)
rules2 = read.PMML("data/rules2.xml")

```

```{r,fig.width=15,fig.height=5}
rules_dt2 <- data.table( lhs = labels( lhs(rules2) ), 
                        rhs = labels( rhs(rules2) ), 
                        quality(rules2) )[ order(-lift), ]
kable(head(rules_dt2,10))
glimpse(rules_dt2)


```

### Rules- Scatter Plot ( For Aisles)

```{r,fig.width=8,fig.height=8}
library(data.table)
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
library(DT)
library(arules)
library(arulesViz)
library(plotly)
library(visNetwork)
library(igraph)

plotly_arules(rules2)

```


```{r message=FALSE, warning=FALSE,include=FALSE}
    subrules2 <- head(sort(rules2, by="confidence"),10)
    ig <- plot( subrules2, method="graph", control=list(type="items") )
    ig_df <- get.data.frame( ig, what = "both" )
    nodesv <- data.frame(
      id = ig_df$vertices$name
      ,value = ig_df$vertices$support # could change to lift or confidence
      ,title = ifelse(ig_df$vertices$label == "",ig_df$vertices$name, ig_df$vertices$label)
      ,ig_df$vertices
    ) 
    
   
    ```
      
      
### Rules- Network Graph Visual (For Aisle)
    ```{r,fig.width=10,fig.height=10}
      ### Network graph visualization Part 2
    
    
    visNetwork(nodes = nodesv, edges =ig_df$edges, width = 900, height = 700) %>%
      visNodes(size = 10) %>%
      visLegend() %>%
      visEdges(smooth = FALSE) %>%
      visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) %>%
      visInteraction(navigationButtons = TRUE) %>%
      visEdges(arrows = 'from') %>%
      visPhysics(
        solver = "barnesHut",
        maxVelocity = 35,
        forceAtlas2Based = list(gravitationalConstant = -6000)
      )
    

```

Recommendations by Implicit feedback-ALS
=====================================  
    
Column {.tabset}
-------------------------------------

### User-Id:1


#### All products recommended to user 1
['Soda', '0% Greek Strained Yogurt', 'Clementines', 'Bag of Organic Bananas', 'Organic Half & Half', 'Apples', 'Zero Calorie Cola', "Crunchy Oats 'n Honey Granola Bars", 'Extra Fancy Unsalted Mixed Nuts', 'Reduced Fat 2% Milk']


### User ID : 38

####All products recommended to user 38

['Organic Hass Avocado', 'Sparkling Water Grapefruit', 'Half & Half', 'Lime Sparkling Water', 'Sparkling Lemon Water', '2% Reduced Fat Milk', 'Organic Yellow Onion', 'Pure Sparkling Water', 'Organic Grape Tomatoes', 'Organic Garlic']



### User ID : 796

####All products recommended to user 796

['Half & Half', 'Organic Avocado', 'Organic Fuji Apple', 'Banana', 'Large Lemon', 'Organic Strawberries', 'Bag of Organic Bananas', 'Unsweetened Almondmilk', 'Raspberries', 'Organic Large Brown Grade AA Cage Free Eggs']


### User ID : 1000

####All products recommended to user 1000

['Spring Water', 'Cucumber Kirby', 'Organic Gala Apples', 'Organic Granny Smith Apple', 'Icelandic Style Skyr Blueberry Non-fat Yogurt', 'Non Fat Raspberry Yogurt', 'Vanilla Skyr Nonfat Yogurt', 'Organic Kiwi', 'Organic Grape Tomatoes', 'Nonfat Icelandic Style Strawberry Yogurt']


### User ID : 421

#### All products recommended to user 421

['Strawberries', 'Organic Yellow Onion', 'Organic Garlic', 'Large Alfresco Eggs', 'Raspberries', 'Organic Garnet Sweet Potato (Yam)', 'Organic Grape Tomatoes', 'Fat Free Milk', 'Icelandic Style Skyr Blueberry Non-fat Yogurt', 'Organic Small Bunch Celery']

